\vspace{-2.5em} Singular value decomposition of multiframe\\ polarized fluorescence microscopes\vspace{-1em}

### 2018-05-15 notes/2018-05-02-multiframe-svd/report/report.tex
Patrick's comments.

%%%%%%%%%%%%
% Document %
%%%%%%%%%%%%
\begin{document}
\title{\vspace{-2.5em} Singular value decomposition of multiframe\\ polarized
  fluorescence microscopes\vspace{-1em}} \author{Talon Chandler, Min Guo, Hari
  Shroff, Rudolf Oldenbourg, Patrick La Rivi\`ere}
\date{\vspace{-1em}\today\vspace{-1em}}
\maketitle
\section{Introduction}
In these notes we will develop the continuous models for several multiframe
polarized fluorescence microscopes. For each design we will calculate the
spatio-angular point spread function, optical transfer function, and singular
system consisting of the singular values, object-space singular functions, and
data-space singular functions.

@@

    [\mathcal{H}f]_n(\rd{}) = \int_{\mbb{S}^2}d\so{}\int_{\mbb{R}^3}d\ro{}\, h_n(\rd{} - \ro{}, \so{})f(\ro{}, \so{}),\qquad  n=1, 2,\ldots,N, \label{eq:fwd}

@@

    [\mathcal{H}^{\dagger}\mb{g}](\ro{}, \so{}) = \sum_{j=1}^N\int_{\mbb{R}^2}d\mb{r}_{d}\, h_n(\rd{} - \ro{}, \so{})g(\rd{}).\label{eq:adj}

@@

  k_{nn'}(\rd{} - \rd{}') &= \int_{\mbb{S}^2}d\so{}\int_{\mbb{R}^2}d\ro{}\, h_n(\rd{} - \ro{}, \so{})h_{n'}(\rd{} - \ro{}, \so{}). \label{eq:kern}

@@

  [\mathcal{H}\mathcal{H}^{\dagger}\mb{v}_{\bs{\rho}, j}]_n(\rd{}) = \mu_{\bs{\rho}, n} \mb{v}_{\bs{\rho}, j}(\rd{}). \label{eq:eigval}

@@

  [\mathcal{H}\mathcal{H}^{\dagger}\mb{v}_{\bs{\rho},j}]_n(\rd{}) = \me{}^{i 2\pi \ro{}\cdot\bs{\rho}} \sum_{n'=1}^N K_{nn'}(\bs{\rho})[\mb{V}_j(\bs{\rho})]_{n'}, \label{eq:exp}

@@

Comparing Eqs. \ref{eq:eigfunc} and \ref{eq:exp} yield the eigenvalue problem

@@

\section{Polarized epi-illumination with unpolarized epi-detection}
We will start by restricting our analysis of epi-illumination microscopes to
in-focus objects. This means that our object space is
$\mbb{L}_2(\mbb{R}^2 \times \mbb{S}^2)$.

\subsection{Point response function}
In the previous notes we showed that the excitation point response function for
polarized epi-illumination is given by
\begin{align}
  h^{\mh{z}}_{\text{exc}}(\so{}; \mh{p}) &= y_0^0(\so{}) - \frac{1}{\sqrt{5}}\tilde{A}y_2^0(\so{}) + \sqrt{\frac{3}{5}}\tilde{B}\left\{[(\mh{p}\cdot\mh{x})^2 - (\mh{p}\cdot\mh{y})^2]y_2^2(\so{}) - 2(\mh{p}\cdot\mh{x})(\mh{p}\cdot\mh{y})y_2^{-2}(\so{})\right\}, \label{eq:genpsf}
\end{align}
where
\begin{subequations}
\begin{align}
  \tilde{A} &\equiv \cos^2(\alpha/2)\cos(\alpha),\\
  \tilde{B} &\equiv \frac{1}{12}(\cos^2\alpha + 4\cos\alpha + 7),
\end{align}\label{eq:coefficients}%
\end{subequations}
and $\alpha \equiv \arcsin(\text{NA}/n_o)$.

We also showed that the point response function for unpolarized epi-detection is
given by
\begin{align}
  h_{\text{det}}(\ro{}, \so{}) &= [{a}^2(r_o) + 2b^2(r_o)]y_0^0(\so{}) + \frac{1}{\sqrt{5}}\left[- a^2(r_o) + 4b^2(r_o)\right]y_2^0(\so{}),
\end{align}
where
\begin{align}
  a(r_o) = \frac{J_1(2\pi \nu_or_o)}{\pi \nu_or_o}, 
  &\hspace{2em}
    b(r_o) = \frac{\text{NA}}{n_o}\left[\frac{J_2(2\pi \nu_or_o)}{\pi \nu_or_o}\right],  \label{eq:abparadef}
  \intertext{and}
  \nu_o \equiv \frac{\text{NA}}{\lambda},&\hspace{2em}
  \text{NA} = n_o\sin\alpha.
\end{align}
The excitation and detection processes are incoherent, so to find the complete
point response function we can multiply the excitation and detection point response
functions which gives
\begin{align}
  h(\ro{}, \so{}; \mh{p}) = h^{\mh{z}}_{\text{exc}}(\so{}; \mh{p})h_{\text{det}}(\ro{}, \so{}) = \frac{1}{\tilde{N}}\sum_{l=0}^{\infty}\sum_{m=-l}^l h_l^m(\ro{}; \mh{p})y_l^m(\so{}), 
\end{align}
where $\tilde{N} = \frac{\tilde{A}}{10} + \frac{1}{2}$ is a normalization constant
and the terms in the series are given by
\begin{align}
  h_0^0(r_o) &= \left[\frac{\tilde{A}}{10} + \frac{1}{2}\right]a^2(r_o) + \left[-\frac{2\tilde{A}}{5} + 1\right]b^2(r_o),\\
  h_2^{-2}(r_o; \mh{p}) &= \left[\frac{9\sqrt{15}}{70}a^2(r_o) + \frac{3\sqrt{15}}{35}b^2(r_o)\right]\tilde{B}[(\mh{p}\cdot\mh{x})^2 - (\mh{p}\cdot\mh{y})^2],\\
  h_2^0(r_o) &= \left[-\frac{\sqrt{5}\tilde{A}}{14} + \frac{\sqrt{5}}{10}\right]a^2(r_o) + \left[-\frac{11\sqrt{5}\tilde{A}}{35} + \frac{2}{\sqrt{5}}\right]b^2(r_o),\\
  h_2^2(r_o; \mh{p}) &= -2\left[\frac{9\sqrt{15}}{70}a^2(r_o) + \frac{3\sqrt{15}}{35}b^2(r_o)\right]\tilde{B}(\mh{p}\cdot\mh{x})(\mh{p}\cdot\mh{y}),\\
  h_4^{-2}(r_o; \mh{p}) &= \left[-\frac{3\sqrt{5}}{70}a^2(r_o) + \frac{6\sqrt{5}}{35}b^2(r_o)\right]\tilde{B}[(\mh{p}\cdot\mh{x})^2 - (\mh{p}\cdot\mh{y})^2],\\
  h_4^0(r_o) &= \frac{3\tilde{A}}{35}[a^2(r_o) - 4b^2(r_o)],\\
  h_4^2(r_o; \mh{p}) &= -2\left[-\frac{3\sqrt{5}}{70}a^2(r_o) + \frac{6\sqrt{5}}{35}b^2(r_o)\right]\tilde{B}(\mh{p}\cdot\mh{x})(\mh{p}\cdot\mh{y}),
\end{align}
and all other $h_l^m$ terms in the series are zero. Finally, the complete
forward model for this class of microscope is given by
\begin{align}
    g_n(\rd{}) = \int_{\mbb{S}^2}d\so{}\int_{\mbb{R}^3}d\ro{}\, h(\rd{} - \ro{}, \so{}; \mh{p}_n)f(\ro{}, \so{}),\qquad  n=1, 2,\ldots,N. \label{eq:fwdpolillum}
\end{align}
Notice that we have used a single function $h(\rd{} - \ro{}, \so{}; \mh{p}_n)$
to describe the complete imaging system, and we have shifted the index $n$ to
the illumination polarization. A typical choice of frames is given by $N=4$ and
\begin{align}
  \mh{p}_1 = \mh{x},\hspace{2em}
  \mh{p}_2 = \frac{1}{\sqrt{2}}(\mh{x} + \mh{y}),\hspace{2em}
  \mh{p}_3 = \mh{y},\hspace{2em}
  \mh{p}_4 = \frac{1}{\sqrt{2}}(\mh{x} - \mh{y}). 
\end{align}
\indent The point response function of the microscope now contains seven terms
in the $l=0$, 2, and 4 bands. Each term is radially symmetric which means that
we expect a radially symmetric point spread function for any angular
distribution of fluorophores. Notice that the $h_l^0$ terms do not depend on the
polarizer orientation, while the $h_l^{-2}$ and $h_l^{2}$ do depend on the
polarizer orientation. This is because the fluorophore distributions
corresponding to the $h_l^0$ terms are rotationally symmetric about the optic
axis, while the other terms are not rotationally symmetric.

\subsection{Optical transfer function}
The optical transfer function for this microscope is given by
\begin{align}
  H_l^m(\nu; \mh{p}) = \frac{1}{\tilde{M}}\sum_{l'=0}^{\infty}\sum_{m'=-l'}^{l'} H_{l'}^{m'}(\nu; \mh{p})\delta(l - l', m - m'),  
\end{align}
where
$\tilde{M} = \left[\frac{\tilde{A}}{10} + \frac{1}{2}\right] + \left(\frac{\text{NA}}{n_o}\right)\left[-\frac{2\tilde{A}}{10} + \frac{1}{2}\right]$ is a
normalization constant, and
\begin{align}
  H_0^0(\nu) &= \left[\frac{\tilde{A}}{10} + \frac{1}{2}\right]A(\nu) + \left[-\frac{2\tilde{A}}{5} + 1\right]B(\nu),\\
  H_2^{-2}(\nu; \mh{p}) &= \left[\frac{9\sqrt{15}}{70}A(\nu) + \frac{3\sqrt{15}}{35}B(\nu)\right]\tilde{B}[(\mh{p}\cdot\mh{x})^2 - (\mh{p}\cdot\mh{y})^2],\\
  H_2^0(\nu) &= \left[-\frac{\sqrt{5}\tilde{A}}{14} + \frac{\sqrt{5}}{10}\right]A(\nu) + \left[-\frac{11\sqrt{5}\tilde{A}}{35} + \frac{2}{\sqrt{5}}\right]B(\nu),\\
  H_2^2(\nu; \mh{p}) &= -2\left[\frac{9\sqrt{15}}{70}A(\nu) + \frac{3\sqrt{15}}{35}B(\nu)\right]\tilde{B}(\mh{p}\cdot\mh{x})(\mh{p}\cdot\mh{y}),\\
  H_4^{-2}(\nu; \mh{p}) &= \left[-\frac{3\sqrt{5}}{70}A(\nu) + \frac{6\sqrt{5}}{35}B(\nu)\right]\tilde{B}[(\mh{p}\cdot\mh{x})^2 - (\mh{p}\cdot\mh{y})^2],\\
  H_4^0(\nu) &= \frac{3\tilde{A}}{35}[A(\nu) - 4B(\nu)],\\
  H_4^2(\nu; \mh{p}) &= -2\left[-\frac{3\sqrt{5}}{70}A(\nu) + \frac{6\sqrt{5}}{35}B(\nu)\right]\tilde{B}(\mh{p}\cdot\mh{x})(\mh{p}\cdot\mh{y}),
\end{align}
all other $H_l^m$ terms in the series are zero, and  
\begin{align}
  A(\nu) &= \frac{2}{\pi}\left[\arccos\left(\frac{\nu}{2\nu_o}\right) - \frac{\nu}{2\nu_o}\sqrt{1 - \left(\frac{\nu}{2\nu_o}\right)^2}\right]\Pi\left(\frac{\nu}{2\nu_o}\right),\\
  B(\nu) &= \frac{1}{\pi}\left(\frac{\text{NA}}{n_o}\right)^2\left[\arccos\left(\frac{\nu}{2\nu_o}\right) - \left[3 - 2\left(\frac{\nu}{2\nu_o}\right)^2\right]\frac{\nu}{2\nu_o} \sqrt{1 - \left(\frac{\nu}{2\nu_o}\right)^2}\right]\Pi\left(\frac{\nu}{2\nu_o}\right).                 
\end{align}

>>>>

%%%%%%%%%%%%
% Document %
%%%%%%%%%%%%
\begin{document}
\title{\vspace{-2.5em} Singular value decomposition of multiframe\\ polarized
  fluorescence microscopes\vspace{-1em}} \author{Talon Chandler, Min Guo, Hari
  Shroff, Rudolf Oldenbourg, Patrick La Rivi\`ere}
\date{\vspace{-1em}\today\vspace{-1em}}
\maketitle
\section{Introduction}
% \textbf{Note:} While writing these notes I realized that multiframe microscopes
% can be analyzed more generally by considering them as sampled versions of
% continuous microscope designs. The results in these notes are correct, but I've
% abandoned these notes to work on the continuous designs.

In these notes we will develop the continuous models for several multiframe
polarized fluorescence microscopes. For each design we will calculate the
spatio-angular point spread function, optical transfer function, and singular
system consisting of the singular values, object-space singular functions, and
data-space singular functions.

@@

    g_n(\rd{}) = [\mathcal{H}f]_n(\rd{}) = \int_{\mbb{S}^2}d\so{}\int_{\mbb{R}^3}d\ro{}\, h_n(\rd{} - \ro{}, \so{})f(\ro{}, \so{}),\qquad  n=1, 2,\ldots,N, \label{eq:fwd}

@@

    [\mathcal{H}^{\dagger}\mb{g}](\ro{}, \so{}) = \sum_{j=1}^N\int_{\mbb{R}^2}d\mb{r}_{d}\, h_n(\rd{} - \ro{}, \so{})g_n(\rd{}).\label{eq:adj}

@@

  k_{nn'}(\rd{} - \rd{}') &= \int_{\mbb{S}^2}d\so{}\int_{\mbb{R}^2}d\ro{}\, h_n(\rd{} - \ro{}, \so{})h_{n'}(\rd{}' - \ro{}, \so{}). \label{eq:kern}

@@

  [\mathcal{H}\mathcal{H}^{\dagger}\mb{v}_{\bs{\rho}, j}]_n(\rd{}) = \mu_{\bs{\rho}, j} [\mb{v}_{\bs{\rho}, j}(\rd{})]_n. \label{eq:eigval}

@@

  [\mathcal{H}\mathcal{H}^{\dagger}\mb{v}_{\bs{\rho},j}]_n(\rd{}) = \me{}^{i 2\pi \rd{}\cdot\bs{\rho}} \sum_{n'=1}^N K_{nn'}(\bs{\rho})[\mb{V}_j(\bs{\rho})]_{n'}, \label{eq:exp}

@@

Comparing Eqs. \ref{eq:eigfunc} and \ref{eq:exp} yields the eigenvalue problem

@@

% \section{Polarized epi-illumination with unpolarized epi-detection}
% We will start by restricting our analysis of epi-illumination microscopes to
% in-focus objects. This means that our object space is
% $\mbb{L}_2(\mbb{R}^2 \times \mbb{S}^2)$.

% \subsection{Point response function}
% In the previous notes we showed that the excitation point response function for
% polarized epi-illumination is given by
% \begin{align}
%   h^{\mh{z}}_{\text{exc}}(\so{}; \mh{p}) &= y_0^0(\so{}) - \frac{1}{\sqrt{5}}\tilde{A}y_2^0(\so{}) + \sqrt{\frac{3}{5}}\tilde{B}\left\{[(\mh{p}\cdot\mh{x})^2 - (\mh{p}\cdot\mh{y})^2]y_2^2(\so{}) - 2(\mh{p}\cdot\mh{x})(\mh{p}\cdot\mh{y})y_2^{-2}(\so{})\right\}, \label{eq:genpsf}
% \end{align}
% where
% \begin{subequations}
% \begin{align}
%   \tilde{A} &\equiv \cos^2(\alpha/2)\cos(\alpha),\\
%   \tilde{B} &\equiv \frac{1}{12}(\cos^2\alpha + 4\cos\alpha + 7),
% \end{align}\label{eq:coefficients}%
% \end{subequations}
% and $\alpha \equiv \arcsin(\text{NA}/n_o)$.

% We also showed that the point response function for unpolarized epi-detection is
% given by
% \begin{align}
%   h_{\text{det}}(\ro{}, \so{}) &= [{a}^2(r_o) + 2b^2(r_o)]y_0^0(\so{}) + \frac{1}{\sqrt{5}}\left[- a^2(r_o) + 4b^2(r_o)\right]y_2^0(\so{}),
% \end{align}
% where
% \begin{align}
%   a(r_o) = \frac{J_1(2\pi \nu_or_o)}{\pi \nu_or_o}, 
%   &\hspace{2em}
%     b(r_o) = \frac{\text{NA}}{n_o}\left[\frac{J_2(2\pi \nu_or_o)}{\pi \nu_or_o}\right],  \label{eq:abparadef}
%   \intertext{and}
%   \nu_o \equiv \frac{\text{NA}}{\lambda},&\hspace{2em}
%   \text{NA} = n_o\sin\alpha.
% \end{align}
% The excitation and detection processes are incoherent, so to find the complete
% point response function we can multiply the excitation and detection point response
% functions which gives
% \begin{align}
%   h(\ro{}, \so{}; \mh{p}) = h^{\mh{z}}_{\text{exc}}(\so{}; \mh{p})h_{\text{det}}(\ro{}, \so{}) = \frac{1}{\tilde{N}}\sum_{l=0}^{\infty}\sum_{m=-l}^l h_l^m(\ro{}; \mh{p})y_l^m(\so{}), 
% \end{align}
% where $\tilde{N} = \frac{\tilde{A}}{10} + \frac{1}{2}$ is a normalization constant
% and the terms in the series are given by
% \begin{align}
%   h_0^0(r_o) &= \left[\frac{\tilde{A}}{10} + \frac{1}{2}\right]a^2(r_o) + \left[-\frac{2\tilde{A}}{5} + 1\right]b^2(r_o),\\
%   h_2^{-2}(r_o; \mh{p}) &= \left[\frac{9\sqrt{15}}{70}a^2(r_o) + \frac{3\sqrt{15}}{35}b^2(r_o)\right]\tilde{B}[(\mh{p}\cdot\mh{x})^2 - (\mh{p}\cdot\mh{y})^2],\\
%   h_2^0(r_o) &= \left[-\frac{\sqrt{5}\tilde{A}}{14} + \frac{\sqrt{5}}{10}\right]a^2(r_o) + \left[-\frac{11\sqrt{5}\tilde{A}}{35} + \frac{2}{\sqrt{5}}\right]b^2(r_o),\\
%   h_2^2(r_o; \mh{p}) &= -2\left[\frac{9\sqrt{15}}{70}a^2(r_o) + \frac{3\sqrt{15}}{35}b^2(r_o)\right]\tilde{B}(\mh{p}\cdot\mh{x})(\mh{p}\cdot\mh{y}),\\
%   h_4^{-2}(r_o; \mh{p}) &= \left[-\frac{3\sqrt{5}}{70}a^2(r_o) + \frac{6\sqrt{5}}{35}b^2(r_o)\right]\tilde{B}[(\mh{p}\cdot\mh{x})^2 - (\mh{p}\cdot\mh{y})^2],\\
%   h_4^0(r_o) &= \frac{3\tilde{A}}{35}[a^2(r_o) - 4b^2(r_o)],\\
%   h_4^2(r_o; \mh{p}) &= -2\left[-\frac{3\sqrt{5}}{70}a^2(r_o) + \frac{6\sqrt{5}}{35}b^2(r_o)\right]\tilde{B}(\mh{p}\cdot\mh{x})(\mh{p}\cdot\mh{y}),
% \end{align}
% and all other $h_l^m$ terms in the series are zero. Finally, the complete
% forward model for this class of microscope is given by
% \begin{align}
%     g_n(\rd{}) = \int_{\mbb{S}^2}d\so{}\int_{\mbb{R}^3}d\ro{}\, h(\rd{} - \ro{}, \so{}; \mh{p}_n)f(\ro{}, \so{}),\qquad  n=1, 2,\ldots,N. \label{eq:fwdpolillum}
% \end{align}
% Notice that we have used a single function $h(\rd{} - \ro{}, \so{}; \mh{p}_n)$
% to describe the complete imaging system, and we have shifted the index $n$ to
% the illumination polarization. A typical choice of frames is given by $N=4$ and
% \begin{align}
%   \mh{p}_1 = \mh{x},\hspace{2em}
%   \mh{p}_2 = \frac{1}{\sqrt{2}}(\mh{x} + \mh{y}),\hspace{2em}
%   \mh{p}_3 = \mh{y},\hspace{2em}
%   \mh{p}_4 = \frac{1}{\sqrt{2}}(\mh{x} - \mh{y}). 
% \end{align}
% \indent The point response function of the microscope now contains seven terms
% in the $l=0$, 2, and 4 bands. Each term is radially symmetric which means that
% we expect a radially symmetric point spread function for any angular
% distribution of fluorophores. Notice that the $h_l^0$ terms do not depend on the
% polarizer orientation, while the $h_l^{-2}$ and $h_l^{2}$ do depend on the
% polarizer orientation. This is because the fluorophore distributions
% corresponding to the $h_l^0$ terms are rotationally symmetric about the optic
% axis, while the other terms are not rotationally symmetric.

% \subsection{Optical transfer function}
% The optical transfer function for this microscope is given by
% \begin{align}
%   H_l^m(\nu; \mh{p}) = \frac{1}{\tilde{M}}\sum_{l'=0}^{\infty}\sum_{m'=-l'}^{l'} H_{l'}^{m'}(\nu; \mh{p})\delta(l - l', m - m'),  
% \end{align}
% where
% $\tilde{M} = \left[\frac{\tilde{A}}{10} + \frac{1}{2}\right] + \left(\frac{\text{NA}}{n_o}\right)\left[-\frac{2\tilde{A}}{10} + \frac{1}{2}\right]$ is a
% normalization constant, and
% \begin{align}
%   H_0^0(\nu) &= \left[\frac{\tilde{A}}{10} + \frac{1}{2}\right]A(\nu) + \left[-\frac{2\tilde{A}}{5} + 1\right]B(\nu),\\
%   H_2^{-2}(\nu; \mh{p}) &= \left[\frac{9\sqrt{15}}{70}A(\nu) + \frac{3\sqrt{15}}{35}B(\nu)\right]\tilde{B}[(\mh{p}\cdot\mh{x})^2 - (\mh{p}\cdot\mh{y})^2],\\
%   H_2^0(\nu) &= \left[-\frac{\sqrt{5}\tilde{A}}{14} + \frac{\sqrt{5}}{10}\right]A(\nu) + \left[-\frac{11\sqrt{5}\tilde{A}}{35} + \frac{2}{\sqrt{5}}\right]B(\nu),\\
%   H_2^2(\nu; \mh{p}) &= -2\left[\frac{9\sqrt{15}}{70}A(\nu) + \frac{3\sqrt{15}}{35}B(\nu)\right]\tilde{B}(\mh{p}\cdot\mh{x})(\mh{p}\cdot\mh{y}),\\
%   H_4^{-2}(\nu; \mh{p}) &= \left[-\frac{3\sqrt{5}}{70}A(\nu) + \frac{6\sqrt{5}}{35}B(\nu)\right]\tilde{B}[(\mh{p}\cdot\mh{x})^2 - (\mh{p}\cdot\mh{y})^2],\\
%   H_4^0(\nu) &= \frac{3\tilde{A}}{35}[A(\nu) - 4B(\nu)],\\
%   H_4^2(\nu; \mh{p}) &= -2\left[-\frac{3\sqrt{5}}{70}A(\nu) + \frac{6\sqrt{5}}{35}B(\nu)\right]\tilde{B}(\mh{p}\cdot\mh{x})(\mh{p}\cdot\mh{y}),
% \end{align}
% all other $H_l^m$ terms in the series are zero, and  
% \begin{align}
%   A(\nu) &= \frac{2}{\pi}\left[\arccos\left(\frac{\nu}{2\nu_o}\right) - \frac{\nu}{2\nu_o}\sqrt{1 - \left(\frac{\nu}{2\nu_o}\right)^2}\right]\Pi\left(\frac{\nu}{2\nu_o}\right),\\
%   B(\nu) &= \frac{1}{\pi}\left(\frac{\text{NA}}{n_o}\right)^2\left[\arccos\left(\frac{\nu}{2\nu_o}\right) - \left[3 - 2\left(\frac{\nu}{2\nu_o}\right)^2\right]\frac{\nu}{2\nu_o} \sqrt{1 - \left(\frac{\nu}{2\nu_o}\right)^2}\right]\Pi\left(\frac{\nu}{2\nu_o}\right).                 
% \end{align}
